---
title: "STA 141 - Fundamentals of Statistical Data Science - Final project"
author: "Mina Sarmah"
params:
  term: Winter 2023
  duedate: '2023-03-24'
  names: Mina Sarmah
  email: msarmah@ucdavis.edu
  total_points: XX
  grade: NA
output: 
   html_document: default
   pdf_document: default
---




# Working with the New York Times COVID-19 Data

As we have seen in class, COVID-19 data collected by the *New York
Times* are available in a repository on Github.  In this final project we
will work with these data to reproduce some of the elements of the
*Times*'s reporting on COVID-19.  

```{r include = FALSE, echo = TRUE} 
diff.eg <- quote(diff(c(1, 3, 6, 10)))
```

# Exercises

1. The file `us.csv` contains aggregated data for the entire U.S.  In
   this file, the `cases` and `deaths` columns represent *cumulative*
   cases and deaths due to covid. The `diff()` function can be used to
   compute the differences between each consecutive element of a
   vector, so it could be used to compute the daily numbers of cases
   and deaths.  However, `diff()` returns a vector of length one less
   than the length of the original vector (e.g., `r noquote(deparse(diff.eg))`
   returns `r eval(diff.eg)`) and this can make it somewhat
   inconvenient to use when transforming columns of data frames.
   
   An alternative is to us the more general `filter()` function with
   an appropriate choice of the `filter` and `sides` arguments.  Thus function
   can also be used to computing running averages and similar quantities.
   
    a. Read the file `us.csv` into R as the data frame `us` and do the
       following:
	  
        * Transform the `date` column into a column of class `Date`.
        * Use `filter()` to add a column named `new_cases` containing the
          number of new cases reported on each date. The first
          value in this column will be `NA`.
        * Use `filter()` to add a column named `new_deaths` containing the
          number of new deaths reported on each date. The first
          value in this column will be `NA`.
        * Use `filter()` to add a column named `avg_new_cases` where
     	    each element represents the mean number of new cases for the
     	    previous 7 days (inclusive of the current day).  The first 7
     	    values in this column will be `NA`.
        * Use `filter()` to add a column named `avg_new_deaths` where
     	    each element represents the mean number of new deaths for the
     	    previous 7 days (inclusive of the current day).  The first 7
     	    values in this column will be `NA`.
			
       Note that the `filter()` function used here is
       `stats::filter()` from the `stats` package, which is loaded by
       default in R. (The `dplyr` package has a completely different
       `filter()` function which plays an important role in the
       "tidyverse".  If you have problems using `filter()`, you should
       make sure that you do NOT have the `dplyr` package loaded.  If
       you do, then you will need to explicitly type out
       `stats::filter()` to get the `stats` version.) 

First we will read in us.csv. Below is the first 10 outputs of the data set. 
```{r, echo=FALSE}
# read us.csv as data frame us
setwd("~/Documents/sta/141a/hw3")
us = read.csv("us.csv")
# transform date column to Date class
# currently a character class 
# use as.Date to convert date
us$date = as.Date(us$date)
head(us,n=10)
```

Next, we will create the new_cases, new_deaths, avg_new_cases, and avg_new_deaths columns and print the first 10 outputs of our updated us data set.
```{r, echo=FALSE}
# use filter to find the number of new cases every day
new_cases = stats::filter(us$cases,c(1,-1), method = "convolution", sides = 1)
new_cases = as.integer(new_cases)
# now make new_cases column
us = cbind(us,new_cases)
# use filter to find the number of new deaths every day 
new_deaths = stats::filter(us$deaths,c(1,-1), method = "convolution", sides = 1)
new_deaths = as.integer(new_deaths)
# now make new_deaths column
us = cbind(us,new_deaths)
# use filter to find the mean number of new cases for the previous 7 days 
avg_new_cases = stats::filter(us$new_cases, rep(1/7,7), method = "convolution", sides = 1)
avg_new_cases = as.integer(avg_new_cases)
# avg_new_cases column
us = cbind(us,avg_new_cases)
# use filter to find the mean number of new deaths for the previous 7 days 
avg_new_deaths = stats::filter(us$new_deaths, rep(1/7,7), method = "convolution", sides = 1)
avg_new_deaths = as.integer(avg_new_deaths)
# avg_new_cases column
us = cbind(us,avg_new_deaths)
head(us, n=10)
```
     	  
    b. Create a plot of daily cases similar to the one found at the
  	   top of [this
  	   page](https://www.nytimes.com/interactive/2020/us/coronavirus-us-cases.html).
  	   Plot only data beginning from 2020-03-01. (Note that this plot
  	   and a similar plot for deaths appear again about 1/3 of the way
  	   down the page.)

       Try to do this using the formula method of the `plot()` function with the
       optional arguments `type = "h"`, `col = "gray"`, and `data = us`
       and using the `subset` argument to plot only the data for dates 2020-03-01
       and after. (You may also wish to experiment with the optional argument `lwd`.)
       
       Then use the formula interface to the `lines()` function to add the curve
       showing the seven-day running average. (Again, you may wish to experiment
       with the optional argument `lwd`.)

```{r, echo=FALSE}
# subset the dates to be after 3/1/2020
us_subset = subset(us,us$date>="2020-03-01", drop=FALSE)
us_subset$date = as.Date(us_subset$date)
options(scipen=99)
# creating the new cases plot
plot(us_subset$date,
     us_subset$new_cases,
     type="h", 
     main = "US New Reported Cases after March 1st, 2020", xlab = "Dates", ylab = "Cases", 
     col="gray")
lines(us_subset$date, us_subset$avg_new_cases, col="red")
```

For the plot of new reported daily cases for the U.S. after March 1, 2020, we can see there are multiple peaks across three years. The first peak in cases we can see happens towards the end of 2020, starting October and decreasing April 2021. This was likely a response to COVID precautions loosening as cases appeared to decline right before the beginning of our first peak. The biggest spike we see in this plot occurs at the beginning of 2022, around January and February. This spike is at least twice as large as the first peak, and has two smaller peaks around October and November 2021 and April through July 2022. The largest peak in cases could have been the result of post-holiday traveling alongside states loosening COVID-19 prevention guidelines. 

    c. Repeat part (b) for deaths.
    
```{r, echo=FALSE}
# subset the dates to be after 3/1/2020
us_subset = subset(us,us$date>="2020-03-01", drop=FALSE)
us_subset$date = as.Date(us_subset$date)
options(scipen=99)
# creating the new deaths plot
plot(us_subset$date,
     us_subset$new_deaths,
     type="h", 
     main = "US New Reported Deaths after March 1st, 2020", xlab = "Dates", ylab = "Deaths", 
     col="gray")
lines(us_subset$date, us_subset$avg_new_deaths, col="green")
```

Our US New Reported Deaths graph has peaks of less intensity compared to our US New Reported Cases plot aside from November 11, 2022. This singular spike will stay unexplained as we do not know the cause of it or if it is an error in the data set. The first peak we see happens during March to May of 2020. Although cases were not as high during this time, many doctors were unsure how to treat COVID-19, and symptoms escalated quickly causing the patients to die. The next two spikes we see happen towards the beginning of 2021 and towards the end of 2021 into 2022. This coincides with our New Reported Cases plot. Another interesting thing to note about this graph is the negative change in new reported deaths occurring as we come out of our second peak. This change indicates that the number of deaths dropped dramatically. Further support for this can be seen as there are no more irregular spikes in newly reported deaths throughout 2022 and 2023.

2. The file `us-states.csv` contains county-level data for the U.S.

    a. Read `us-states.csv` into R as the data frame `us_states`
       and transform the date column into a column of class `Date`.

First we will read in us-states.csv. Below is the first 10 outputs of the data set. 
```{r, echo=FALSE}
# read in us-states
setwd("~/Documents/sta/141A/hw3")
us_states = read.csv("us-states.csv")
# transform to date class
us_states$date = as.Date(us_states$date)
head(us_states, n=10)
```
	   
    b. Use `subset()` to extract the data for the state of California
       and save it as a data frame named `California`. Be sure that the
	   rows are correctly ordered by date, and then repeat parts 1b
	   and 1c of this assignment for California, i.e., plot the number of
	   daily new cases and deaths, along with their 7-day running averages.

Next, we will create the new_cases, new_deaths, avg_new_cases, and avg_new_deaths columns and print the first 10 outputs of our updated us-states data set.
```{r, echo=FALSE}
# make California subset
California = subset(us_states,us_states$state=="California", drop=FALSE)
California$date = as.Date(California$date)
# create new cases for subset
new_cases = stats::filter(California$cases,c(1,-1), method = "convolution", sides = 1)
new_cases = as.integer(new_cases)
California = cbind(California,new_cases)
# create new deaths for subset
new_deaths = stats::filter(California$deaths,c(1,-1), method = "convolution", sides = 1)
new_deaths = as.integer(new_deaths)
California = cbind(California,new_deaths)
# create average new cases for subset
avg_new_cases = stats::filter(California$new_cases, rep(1/7,7), method = "convolution", sides = 1)
avg_new_cases = as.integer(avg_new_cases)
California = cbind(California,avg_new_cases)
# create average new deaths for subset
avg_new_deaths = stats::filter(California$new_deaths, rep(1/7,7), method = "convolution", sides = 1)
avg_new_deaths = as.integer(avg_new_deaths)
California = cbind(California,avg_new_deaths)
# now make new_cases, new_deaths, avg_new_cases, avg_new_deaths column
head(California,n=10)
```

```{r, echo=FALSE}
# creating the new cases plot
plot(California$date,
     California$new_cases,
     type="h", 
     main = "California New Reported Cases", xlab = "Dates", ylab = "Cases", 
     col="gray")
lines(California$date, California$avg_new_cases, col="blue")
```

The plot above mirrors the new reported cases plot in California for the entirety of the US. Similarly, the first spike in cases we see occurs during the end of 2020 and into 2021 with the second major peak occurring towards the beginning of 2022. In this plot and in the plot for the entire US, it appears that in times of peak travel and colder weather, cases in COVID-19 increase far more compared to times of low travel and warmer weather.

```{r, echo=FALSE}
# creating the new deaths plot
plot(California$date,
     California$new_deaths,
     type="h", 
     main = "California New Reported Deaths", xlab = "Dates", ylab = "Deaths", 
     col="gray")
lines(California$date, California$avg_new_deaths, col="orange")
```

From our plot above we can see California has one major peak in new reported deaths during the beginning of 2021, aside from November 11, 2022. This increase in new reported deaths parallels the first peak in cases we saw in the California New Cases plot. As we come out of this peak, we also see a few days through the summer months of 2021 where there are negative changes in new reported deaths.

3. The file `us-counties.csv` contains county-level data for the U.S.

    a. Read `us-counties.csv` into R as the data frame `us_counties`
       and transform the date column into a column of class `Date`.

First we will read in us-counties.csv. Below is the first 10 outputs of the data set. 
```{r, echo=FALSE}
# read in us-counties
setwd("~/Documents/sta/141A/hw3")
us_counties = read.csv("us-counties.csv")
# transform date to Date class
us_counties$date = as.Date(us_counties$date)
head(us_counties, n=10)
```

    b. Use `subset()` to extract the data for Yolo County, California,
       and save it as a data frame named `Yolo`. Be sure that the
	   rows are correctly ordered by date, and then repeat part 1b
	   this assignment for Yolo County, i.e., plot the number of daily new
	   cases along with their 7-day running average. 

Next, we will create the new_cases, new_deaths, avg_new_cases, and avg_new_deaths columns and print the first 10 outputs of our updated us-counties data set.	   
```{r, echo=FALSE}
# make Yolo subset
Yolo = subset(us_counties,us_counties$county=="Yolo", drop=FALSE)
Yolo$date = as.Date(Yolo$date)
# create new cases for subset
new_cases = stats::filter(Yolo$cases,c(1,-1), method = "convolution", sides = 1)
new_cases = as.integer(new_cases)
Yolo = cbind(Yolo,new_cases)
# create new deaths for subset
new_deaths = stats::filter(Yolo$deaths,c(1,-1), method = "convolution", sides = 1)
new_deaths = as.integer(new_deaths)
Yolo = cbind(Yolo,new_deaths)
# create average new cases for subset
avg_new_cases = stats::filter(Yolo$new_cases, rep(1/7,7), method = "convolution", sides = 1)
avg_new_cases = as.integer(avg_new_cases)
Yolo = cbind(Yolo,avg_new_cases)
# create average new deaths for subset
avg_new_deaths = stats::filter(Yolo$new_deaths, rep(1/7,7), method = "convolution", sides = 1)
avg_new_deaths = as.integer(avg_new_deaths)
Yolo = cbind(Yolo,avg_new_deaths)
head(Yolo,n=10)
```

```{r, echo=FALSE}
# creating Yolo new cases plot
plot(Yolo$date,
     Yolo$new_cases,
     type="h", 
     main = "Yolo County New Reported Cases", xlab = "Dates", ylab = "Cases", 
     col="gray")
lines(Yolo$date, Yolo$avg_new_cases, col="pink")
```

For our Yolo county new reported cases plot, we see that the first peak in cases appears around late 2020 and early 2021; the second wave spikes during the beginning months of 2022. The second peak in 2022 is much larger than our first peak, and a main factor of that could be the transition of students returning to their educational institutions (ex. university, high school) after winter break. Coupled with the mandatory testing of all UC Davis students and staff after winter break, many cases of COVID were reported and thus the peak in cases occurred.

	   Q: What do you notice when comparing the plot of daily new cases
       in Yolo county to the analogous plot for the state of
	   California as a whole?  What might explain what you are seeing?

When comparing the plot of daily new cases in Yolo county to the state of California as a whole, we can see that the waves occur at the same time but with less intensity in Yolo county. The first peak we see towards the end of 2020 and beginning of 2021 is likely due to people traveling during the holiday season for the first time after the rise of COVID-19. This movement of people across state and country borders freely is likely a leading cause of the strongest wave of COVID-19 cases in the beginning of 2022.  More specifically to Yolo County,  UC Davis students and staff come from all over the country and beyond; the spread of COVID-19 was an expected risk of winter break.  